---
title: "placeR_vignette"
author: "Ulrich Matter"
date: "January 17, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set Up

Install and load the package:
```{r}
#devtools::install_github("umatter/placeR")
#library(placeR)

devtools::load_all()
```

The package's request functions expect a variable `api.key` with the Google Places API Key stored in it in the global environment.
```{r}
api.key <- readLines("_misc/key")[1]
```


## Example of simple searches

### Search Restaurants in New York

Search restaurants in a given radius around specific coordinates with `searchNearby()`.
```{r}
nyrest <- searchNearby(location="40.7,-74.0", radius=5000, types="restaurant")
```

Inspect the results. The formatted data can be extracted via `placesData()`. 
```{r}
# check the class of the returned object
class(nyrest)

# extract and inspect data
ny <- placesData(nyrest)
dim(ny)
str(ny)
head[1:3,1:4]
```

Or we can plot the results via the plot method for objects of class "`placesearch`":
```{r}
plot(nyrest, ggmap.zoom=12)
```

## Large area searches

### Detailed example, step by step

Define the search parameters. Two points (coordinates) define the bounding box.
```{r}
sa <- c(5.5, 11, 45.5, 48)
a.df <- as.data.frame(matrix(data=sa, nrow=2, ncol=2))
names(a.df) <- c("x","y")
```

Set the radius of each individual radar search
```{r}
r=50000
```

Compute search points for searchRadar. That is, at which points of the overall area should a radar search be executed.
```{r}
sdlpoints <- getPointsArea(area=a.df, radius=r)
sdllocs <- locations(sdlpoints)
```

Now, we execute the radar search for the first five locations:
```{r}
sdltest <- searchRadar(location=sdllocs[1:5], types="restaurant", saveDL=FALSE)
```

Extract the references to each place/location for the detail search, and execute the details search
```{r}

# Not yet implemented
# refstest <- na.omit(as.character(sdltest$reference))
# # search details to the references
# details_test <- placeDetails(reference=refstest[1:150],
#                              saveDL=TRUE, 
#                              chunksize=10) 

```


## Example using AreaScan (high level function)
```{r}
# example of search: scan all switzerland
library(sp)
cantons <- readRDS(url("https://biogeo.ucdavis.edu/data/gadm3.6/Rsp/gadm36_CHE_1_sp.rds")) # get a shapefile with swiss admin. borders
print(cantons)

# extract map edges
edges <- getEdges(x=cantons) 

# plot search area
area <- getPointsArea(edges, radius=20000)
plot(area)
```


## Extended example to demonstrate areaScan
The goal of the exercise is to get information on 'all' restaurants in a canton.

First, load the respective shapefile
```{r}
cantons <- readRDS(url("https://biogeo.ucdavis.edu/data/gadm3.6/Rsp/gadm36_CHE_1_sp.rds")) # get a shapefile with swiss admin. borders
ag <- cantons[cantons$NAME_1=="Aargau",] 
bs <- cantons[cantons$NAME_1=="Basel-Stadt",] 
```

Run the area scan. Note that due to the depriciation of the API's radar search function, the results are restricted to 20 (60 if going to next page).
```{r}
# extract map edges
edges <- getEdges(x=ag) 
# run the scan
scanresp <- areaScan(area=edges, radius=8000, types="restaurant", details=FALSE )
```

Extract, clean, and have a look at the data...
```{r}
data <- placesData(scanresp)
data <- data[!duplicated(data$id),]
View(data)
```





